Update of the ozone dry deposition in the OsloCTM3 v1.0

### 2018-11-13 DryDep.tex
A lot about sensitivity studies and some new plots.

In Section~\ref{sec:model_des}, we will briefly describe the OsloCTM3, give a detailed account of the new dry deposition scheme (Section~\ref{subsec:DryDep}) as well as present pre-processing of meteorological input data to compute necessary input to the dry deposition scheme such as begin and duration of greening season (GDAY, GLEN) and photosynthetic photon flux density (PPFD) (Section~\ref{subsec:pre-pro}). In Section~\ref{sec:eval}, we present sensitivity tests with respect to a manifold of parameters in the dry deposition scheme (Section~\ref{subsec:sens}) and validate our results with results from the multi-model intercomparison of \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and to observations both at the surface as well as in the free troposphere (Section~\ref{subsec:obs}). In Section~\ref{sec:disc}, we will summarize our results and draw conclusions for further development of the model (Section~\ref{sec:conc}).
{\bf TODO: Precise outlook on paper when writing is done.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Model description}
\label{sec:model_des}
The OsloCTM3 is a three dimensional global chemistry transport model (CTM). The key components of the OsloCTM3 have been described and evaluated by \citet{GMD:Sovde2012}. A detailed account of the capabilities of the OsloCTM3 in simulating anthropogenic aerosol forcing in the past and recent past using the Community Emission Data System (CEDS) historical emission inventory \citep{GMD:Hoesly2018} is given by \citet{GMD:Lund2018}. All relevant information and papers can also be found on the \href{http://folk.uio.no/mariantl/osloctm3/}{OsloCTM3 webpage}. The OsloCTM3 can also be coupled to the Model of Emissions of Gases and Aerosols from Nature (MEGAN) \citep{ACP:Guenther2006}. A publication focusing on this is planed. Parallelization of the OsloCTM3 has not been optimized as of now. It is possible to run the model in resolution of at least T159N80L60, but it is not very time and memory efficient. Therefore, we run with a reduced horizontal resolution of $2.25^\circ\times2.25^\circ$. In the following, we will give a detailed account of the new dry deposition scheme. Although some of the equations in this section may be \emph{textbook knowledge}, for the sake of completeness, it is important to summarize them, non the less.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ozone dry deposition scheme}
\label{subsec:DryDep}
In the original dry deposition scheme of the OsloCTM3, the state of the atmosphere was not taken into account. Dry deposition velocities were rather parameterized following the work of \citet{AE:Wesely1989} with parameter updates from \citet{JGR:Hough1991}. This means that seasonal day and night average deposition velocities for different land surface types (water, forest, grass, tundra/desert, and ice and snow) were in use. Day was distinguished from night by solar zenith angles below 90\,\unit{^\circ}. Winter was defined by temperatures below 273.15\,\unit{K} for gridboxes containing land masses. For ocean, winter and summer parameters are equal in this parameterization, therefore no distinctive treatment is needed for ocean gridboxes. In addition, a reduced uptake due to snow cover above 1\,\unit{m} for forest and 10\,\unit{cm} for grass/tundra, respectively, was taken into account. We will refer to this parameterization as \emph{Wesely scheme}.\\

@@

As reference height $z$, we use the height at mid-level of the lowermost model level (roughly 8\,\unit{m}), while $u_z$ and $u_*$ are available from the meteorological input fields ($u_* = \sqrt{\frac{\tau_0}{\rho}}$, with $\tau^2 = \text{NSSS}^2 + \text{EWSS}^2$, and Northward turbulent surface stress (NSSS), eastward turbulent surface stress (EWSS)).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Stomatal conductance}
The stomatal conductance is a measure of the rate of \chem{CO_2} entering or evapotranspiration through the stomata of a leaf. There are several environmental conditions affecting the opening and closing of the stomata and hence their capability of respiration (e.g., light, available water, etc.). Stomata sluggishness, a state in which the stomata can no longer fully close, has been reported as ozone induced damage \citep{SR:Hoshika2015}. To reflect part of the underlying mechanism, the stomatal conductance $g_\text{STO}$ in the EMAP scheme is computed using a multiplicative ansatz that is also explained in much detail in \citet{ICP:MappingManual2017}:
%
\begin{equation}
  g_\text{STO} = g^\text{m}_\text{max} \cdot f_\text{phen} \cdot f_\text{light} \cdot \max{\left\{f_\text{min}, f_T \cdot f_D \cdot f_\text{SW}\right\}}.
  \label{eq:stomatal}
\end{equation}
%
The factors herein are normalized and vary within the range $0-1$. They account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_T$), water vapor pressure deficit ($f_D$), and soil water content ($f_\text{SW}$).

@@

The non-stomatal conductance for \chem{O_3} consists of two terms, one depending on vegetation type and one depending on the soil/surface. For a land surface type $N$, we can write
\begin{equation}
  G_\text{ns}^\chem{O_3}(N) = \frac{\text{SAI}(N)}{r_\text{ext}} + \frac{1}{R_\text{inc}(N) + R_\text{gs}^\chem{O_3}(N)}.
  \label{eq:non-stomata-cond}
\end{equation}
$\text{SAI}(N)$ is the surface area index for vegetation type $N$, which is $\text{LAI}$ plus some value representing cuticles and others surfaces. The external leaf resistance is defined by
\begin{equation}
  r_\text{ext} = 2000\,\unit{s\,m^{-1}}\cdot F_T.
\end{equation}
Herein $F_T$ is a temperature correction factor for temperatures below $-1\,\unit{^\circ C}$ and $\{F_T \in \mathbb{R} | (1 \leq F_T \leq 2)\}$
\begin{equation}
  F_T = \exp{(-2\cdot(1+\theta_0))}.
\end{equation}
$\theta_0$ is the surface or 2\,\unit{m} temperature in \unit{^\circ C}. In general, $\text{SAI} \equiv \text{LAI}$ for all land types, except for: 
%
\begin{equation}
  \text{SAI} = 
  \begin{cases}
    \text{LAI}+1 & \emph{if} \quad \text{\emph{forest / wetland}},\\
    \text{LAI}\cdot 5/3.5 & \emph{if} \quad \text{\emph{cropland}, I. part of growing season},\\
    \text{LAI} + 1.5 & \emph{if} \quad \text{\emph{cropland}, II. part of growing season},\\
    0 & \emph{if} \quad \text{\emph{cropland}, winter}.
  \end{cases}
  \label{eq:SAI}
\end{equation}
%
Extending the EMEP scheme to the southern hemisphere, the used growing season for crops is defined in Table~\ref{tab:growing_season}.
\begin{table}[t]
  \caption{Definition of growing season for crops used in the OsloCTM3 in northern hemisphere (NH) and southern hemisphere (SH).}
  \begin{tabular}{lrr}
    \tophline
    & I. part & II. part \\
    & (days) & (days) \\
    \middlehline
    NH & 90--140 & 141--270 \\
    SH & 272--322 & 323--452 \\
    \bottomhline
  \end{tabular}
  %\belowtable{} % Table Footnotes
  \label{tab:growing_season}
\end{table}
%

In this way, vegetation affects the conductance also by being there not only by uptake through the stomata. The in canopy resistance $R_\text{inc}$ is defined for each vegetated land type $N$ as
\begin{equation}
  R_\text{inc} = b \cdot \text{SAI}(N) \cdot \frac{h(N, \text{lat})}{u_*},
\end{equation}
where $h(N, \text{lat})$ is the latitude dependent vegetation height (see explanation at the end of this section) and $b = 14\,\unit{s^{-1}}$ is an empirical constant. According to \citet{ACP:Simpson2012}, the canopy resistance in EMEP does not take temperature and snow into account and is zero for non-vegetated surfaces.\\
$R_\text{gs}^\chem{O_3}(N)$ is tabulated and corrected for temperature by $F_T$ and snow cover fraction $f_\text{snow}$:
\begin{equation}
  \frac{1}{R_\text{gs}^\chem{O_3}(N)} = \frac{1-f_\text{snow}}{\hat{R}_\text{gs}^\chem{O_3}(N)} + \frac{f_\text{snow}}{R_\text{snow}^\chem{O_3}}.
\end{equation}
As initially mentioned, the necessary depth of snow to cover a certain type of vegetation, differs. Therefore, we calculate a snow cover fraction $f_\text{snow}$ using the snow depth $S_D$, which is available in units of meter of water equivalent from the meteorological input data, scaled to 10\,\unit{\%} of the vegetation height.

@@

\subsubsection*{Latitude dependent vegetation height}
The vegetation height $h(N, \text{lat})$ in the EMEP scheme is linearly decreasing with latitude between 60 and 74\,\unit{^\circ north}. To adapt this to a global model, we need to make a few additional assumptions. The tabulated height for each vegetation type $h(N)$ in the EMEP scheme is regarded as constant at mid latitudes (40--60\,\unit{^\circ}). Towards the pol, we decrease the height of each vegetation type using the same rate as described in \citet{ACP:Simpson2012}. At a latitude of 74\,\unit{^\circ} a minimum height of $3/10 \cdot h(N)$ is reached and kept constant. Towards the equator, we increase the height linearly so that at a latitude of 10\,\unit{^\circ} a maximum height of $2 \cdot h(N)$ is achieved which is then held constant. We also assume symmetry in both hemispheres. Presume a typical tree height of 20\,\unit{m} at mid latitudes, this step-wise function yields a height of 8\,\unit{m} at high latitudes and 40\,\unit{m} in the tropics which is not unrealistic. For four example PFTs, results are shown in the Supplement (Fig.~S.2).

@@

In agriculture, there are different empirical rules to estimate the SGS and EGS. The simplest assumption is that greening starts after 5 consecutive days with a daily average temperature above 5\,\unit{^\circ C} and vice versa for EGS. Based on this rule of thumb, we have pre-processed our meteorological input data offline. If, within these 5 days, the temperature drops below / rises above 5\,\unit{^\circ C}, the counters are reset, respectively. First, we used the 5\,\unit{^\circ C}-days criteria for $45\,\unit{^\circ} < \text{lat} < 85\,\unit{^\circ}$ in the NH, but extended it also to $35\,\unit{^\circ} < \text{lat} < 65\,\unit{^\circ}$ in the SH. In all other cases and where the 5\,\unit{^\circ C}-days criteria fails, we use the SMOKE-BEIS parameterization. The described algorithm written in python 2.7 has been included as Supplement~S.4.1. An example map of the computed GLEN using the 5\,\unit{^\circ C}-days criteria in both hemispheres is shown in Fig.~\ref{fig:glen_2015_she}.
%
\begin{figure}[t]
  \includegraphics[width=8.3cm, clip, trim={0.cm 1.75cm 0.cm 1.75cm}]{pictures/GLEN_2005_she.pdf}
  \caption{Pre-processing of greening season from meteorological surface temperature fields. Shown is the total length of the greening season (GLEN) for the year 2005. The 5\,\unit{^\circ C}-days criteria has been used in both hemispheres mid--high latitudes. Ocean has been shaded to indicate that greening season will only affect land.}
  \label{fig:glen_2015_she}
\end{figure}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Photosynthetic photon flux density}
From OpenIFS an accumulated PAR is available, but for practical use in Eq.~(\ref{eq:flight}) it needs to be deaccumulated. Since this field is already integrated over the spectrum $(400-700)\,\unit{nm}$, it is not PAR but PPFD. Therefor, we will refer to the deaccumulated field as PPFD. An example PAR for January 2nd 2005 is shown in Supplement~S.5.1. Deaccumulation of hour 3 till 21, simply means
\begin{equation}
  \text{PPFD}(t_i) = \text{PAR}(t_{i+1})-\text{PAR}(t_i).
\end{equation}
If everything was right in the OpenIFS output, we would not need to do any further processing, however the first timestep of each day has already accumulated PAR from some other date and time in addition to the actual hour 0 data (see Supplement~S.5.2), so it is impossible to accurately deaccumulate this time step. The best choice is subtracting the difference between hour 21 and 12 of the previous day (Supplement~S.5.3)
\begin{equation}
  \text{PPFD}(t=0) = \text{PAR}(t=0) - \left[\text{PAR}(t=21\,\unit{h}-1\,\unit{day})-\text{PAR}(=12\,\unit{h}-1\,\unit{day})\right].
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%      
\section{Evaluation}
\label{sec:eval}
In this section, we present results from a manifold of OsloCTM3 model runs testing different parameters of the EMEP scheme. We focus on the change in total ozone dry deposition, ozone dry deposition velocities, and surface ozone concentration related to the stomatal conductance part of the surface resistance. We evaluate the results in the context of the multi-model comparison and observations
in\citet{ACP:Hardacre2015} (Section~\ref{subsec:model}-\ref{subsec:obs}). The OsloCTM3 is driven by meteorological input fields from \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF - OpenIFS version cy38r1nc4}. CEDS historical emission inventory is used for anthropogenic emissions, while biomass burning is covered by NASA's Global Fire Emissions Database, Version4 (GFEDv4). In the following, we will discuss the various model sensitivity studies.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Sensitivity tests}
\label{subsec:sens}
A summary of specifications of all simulations discussed in this section is given in Table~\ref{tab:simsum}. For simplicity, we only list the tested parameters. An x in the table denotes that the model was run exactly in the configuration as has been decribed in Section~\ref{sec:model_des}. With respect to the sensitivity on light (PPFD), $f_\text{light}\equiv 1$ has been tested. Regarding the phenology, we have tested three different options: No phenology ($f_\text{phen}\equiv 1$), phenology including / excluding the Tropics (TR). Diverging from \citet{ACP:Simpson2012}, the simulations inititally ran with the uppermost soil water level (SWVL1). One sensitivity test was conducted using SWVL4. The meteorological year used in all simulations is 2005. Most of the simulations were run with CEDS emissions for 2014. One test has been conducted to look at the impact of the emissions.
\begin{table*}[t]
  \caption{Summary of specifications of all simulations discussed in this section. For simplicity, only the tested parameters are listed. An x denotes that the model was run exactly in the configuration as has been decribed in Section~\ref{sec:model_des}.}
  \begin{tabular}{lccccccccc}
    \tophline
    Simulation & Wesely scheme & \multicolumn{3}{c}{EMEP scheme} & \multicolumn{2}{c}{Greening season} & $v^\chem{O_3}_\text{ice/snow}$ & Forcings\\
    & & $f_\text{phen}$ & $f_\text{light}$ & $f_\text{SW}$ & SMOKE-BEIS & 5\,\unit{^\circ C}-days & (\unit{cm\,s^{-1}}) & (year)\\
    \middlehline
    Wesely type       & x & \multicolumn{3}{c}{n/a}  & x &       & 1/2000  & 2014 \\
    EMEP\_full        &   & incl. TR & x & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_offLight    &   & incl. TR & 1 & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_offPhen     &   & 1        & x & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_SWVL4       &   & incl. TR & x & SWLV4     & x &       & 1/2000  & 2014 \\
    EMEP\_ppgs        &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2014 \\
    EMEP\_ppgssh      &   & excl. TR & x & SWLV1     &   & NH+SH & 1/2000  & 2014 \\
    EMEP\_ppgssh\_ice &   & excl. TR & x & SWLV1     &   & NH+SH & 1/10000 & 2014 \\
    EMEP\_ppgs\_2005  &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2005 \\
    \bottomhline
  \end{tabular}
  \label{tab:simsum}
  %\belowtable{} % Table Footnotes
\end{table*}
\citep[e.g., $v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{cm\,s^{-1}}$,][]{ACP:Helmig2007}
dry dep velocity over water \citep{JGR:Helmig2012} Impact of ocean? Although it is a low number there is 2\/3 ocean on earth... (Think, I read this in Hardacre paper... -> Luhar)

>>>>

In Section~\ref{sec:model_des}, we will briefly describe the OsloCTM3, give a detailed account of the new dry deposition scheme (Section~\ref{subsec:DryDep}) as well as present pre-processing of meteorological input data to compute necessary input to the dry deposition scheme such as begin and duration of greening season (GDAY, GLEN) and photosynthetic photon flux density (PPFD) (Section~\ref{subsec:pre-pro}). In Section~\ref{sec:eval}, we present sensitivity tests with respect to a manifold of parameters in the dry deposition scheme (Section~\ref{subsec:sens}) and validate our results with results from the multi-model intercomparison of \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and to observations both at the surface as well as in the free troposphere (Section~\ref{subsec:obs}). In Section~\ref{sec:disc}, we will summarize our results and draw conclusions for further development of the model (Section~\ref{sec:conc}).
{\bf TODO: Precise outlook on paper when writing is done.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Model description}
\label{sec:model_des}
The OsloCTM3 is a three dimensional global chemistry transport model (CTM). The key components of the OsloCTM3 have been described and evaluated by \citet{GMD:Sovde2012}. A detailed account of the capabilities of the OsloCTM3 in simulating anthropogenic aerosol forcing in the past and recent past using the Community Emission Data System (CEDS) historical emission inventory \citep{GMD:Hoesly2018} is given by \citet{GMD:Lund2018}. The OsloCTM3 can also be coupled to the Model of Emissions of Gases and Aerosols from Nature (MEGAN v2.10) \citep{ACP:Guenther2006}. A publication focusing on this is planed. Parallelization of the OsloCTM3 has not been optimized as of now. While the meteorological data driving the OsloCTM3 is given in a resolution of at T159N80L60, it is very time and memory consuming to run the CTM with full chemistry at this resolution. Therefore, we run with a reduced horizontal resolution of $2.25^\circ\times2.25^\circ$. In the following, we will give a detailed account of the new dry deposition scheme. Although some of the equations in this section may be \emph{textbook knowledge}, for the sake of completeness, it is important to summarize them, nonetheless.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ozone dry deposition scheme}
\label{subsec:DryDep}
In the original dry deposition scheme of the OsloCTM3, the state of the atmosphere was not taken into account. Dry deposition velocities were rather parameterized following the work of \citet{AE:Wesely1989} with parameter updates from \citet{JGR:Hough1991}. This means that seasonal day and night average deposition velocities for different land surface types (water, forest, grass, tundra/desert, and ice and snow) were in use. Day was distinguished from night by solar zenith angles below 90\,\unit{^\circ}. Winter was defined by temperatures below 273.15\,\unit{K} for gridboxes containing land masses. For ocean, winter and summer parameters are equal in this parameterization, therefore no distinctive treatment is needed for ocean gridboxes. In addition, a reduced uptake due to snow cover above 1\,\unit{m} for forest and 10\,\unit{cm} for grass/tundra, respectively, was taken into account. We will refer to this parameterization as \emph{Wesely scheme}.\\

@@

As reference height $z$, we use the height at mid-level of the lowermost model level (roughly 8\,\unit{m}), while $u_z$ and $u_*$ are available from the meteorological input fields ($u_* = \sqrt{\frac{\tau_0}{\rho}}$, with the turbulent surface stress $\tau_0^2 = \text{NSSS}^2 + \text{EWSS}^2$, and Northward turbulent surface stress (NSSS), eastward turbulent surface stress (EWSS)).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Stomatal conductance}
The stomatal conductance is a measure of the rate of \chem{CO_2} evapotranspiration through the stomata of a leaf. There are several environmental conditions affecting the opening and closing of the stomata and hence their capability of respiration (e.g., light, available water, etc.). Stomata sluggishness, a state in which the stomata can no longer fully close, has been reported as ozone induced damage \citep{SR:Hoshika2015}. To reflect part of the underlying mechanism, the stomatal conductance $g_\text{STO}$ in the EMAP scheme is computed using a multiplicative ansatz that is also explained in much detail in \citet{ICP:MappingManual2017}:
%
\begin{equation}
  g_\text{STO} = g^\text{m}_\text{max} \cdot f_\text{phen} \cdot f_\text{light} \cdot \max{\left\{f_\text{min}, f_T \cdot f_D \cdot f_\text{SW}\right\}}.
  \label{eq:stomatal}
\end{equation}
%
The factors herein are normalized and vary within the range $0-1$. They account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_T$), water vapor pressure deficit ($f_D$), and soil water content ($f_\text{SW}$).

@@

The non-stomatal conductance for \chem{O_3} consists of two terms, one depending on vegetation type and one depending on the soil/surface. For a land surface type $N$, we can write
\begin{equation}
  G_\text{ns}^\chem{O_3}(N) = \frac{\text{SAI}(N)}{r_\text{ext}} + \frac{1}{R_\text{inc}(N) + R_\text{gs}^\chem{O_3}(N)}.
  \label{eq:non-stomata-cond}
\end{equation}
$\text{SAI}(N)$ is the surface area index for vegetation type $N$, which is $\text{LAI}$ plus some value representing cuticles and others surfaces. The external leaf resistance is defined by
\begin{equation}
  r_\text{ext} = 2000\,\unit{s\,m^{-1}}\cdot F_T.
\end{equation}
Herein $F_T$ is a temperature correction factor for temperatures below $-1\,\unit{^\circ C}$ and $\{F_T \in \mathbb{R} | (1 \leq F_T \leq 2)\}$
\begin{equation}
  F_T = \exp{(-2\cdot(1+\theta_0))}.
\end{equation}
$\theta_0$ is the surface or 2\,\unit{m} temperature in \unit{^\circ C}. In general, $\text{SAI} \equiv \text{LAI}$ for all land types, except for: 
%
\begin{equation}
  \text{SAI} = 
  \begin{cases}
    \text{LAI}+1 & \emph{if} \quad \text{\emph{forest / wetland}},\\
    \text{LAI}\cdot 5/3.5 & \emph{if} \quad \text{\emph{cropland}, I. part of growing season},\\
    \text{LAI} + 1.5 & \emph{if} \quad \text{\emph{cropland}, II. part of growing season},\\
    0 & \emph{if} \quad \text{\emph{cropland}, winter}.
  \end{cases}
  \label{eq:SAI}
\end{equation}
%
Extending the EMEP scheme to the southern hemisphere, the used growing season for crops is defined in Table~\ref{tab:growing_season}.
%
\begin{table}[t]
  \caption{Definition of growing season for crops used in the OsloCTM3 in northern hemisphere (NH) and southern hemisphere (SH).}
  \begin{tabular}{lrr}
    \tophline
    & I. part & II. part \\
    & (days) & (days) \\
    \middlehline
    NH & 90--140 & 141--270 \\
    SH & 272--322 & 323--452 \\
    \bottomhline
  \end{tabular}
  %\belowtable{} % Table Footnotes
  \label{tab:growing_season}
\end{table}
%

In this way, vegetation affects the conductance also by being there, not only by uptake through the stomata. The in-canopy resistance $R_\text{inc}$ is defined for each vegetated land type $N$ as
\begin{equation}
  R_\text{inc} = b \cdot \text{SAI}(N) \cdot \frac{h(N, \text{lat})}{u_*},
\end{equation}
where $h(N, \text{lat})$ is the latitude dependent vegetation height (see explanation at the end of this section) and $b = 14\,\unit{s^{-1}}$ is an empirical constant. According to \citet{ACP:Simpson2012}, the canopy resistance in EMEP does not take temperature and snow into account and is zero for non-vegetated surfaces.\\
$R_\text{gs}^\chem{O_3}(N)$ is tabulated and corrected for temperature by $F_T$ and snow cover fraction $f_\text{snow}$:
\begin{equation}
  \frac{1}{R_\text{gs}^\chem{O_3}(N)} = \frac{1-f_\text{snow}}{\hat{R}_\text{gs}^\chem{O_3}(N)} + \frac{f_\text{snow}}{R_\text{snow}^\chem{O_3}}.
\end{equation}
As initially mentioned, the necessary depth of snow to cover a certain type of vegetation, differs. Therefore, we calculate a snow cover fraction $f_\text{snow}$ using the snow depth $S_D$, which is available in units of meter of water equivalent from the meteorological input data, scaled to 10\,\unit{\%} of the vegetation height.

@@

\subsubsection*{Latitude dependent vegetation height}
The vegetation height $h(N, \text{lat})$ in the EMEP scheme is linearly decreasing with latitude between 60 and 74\,\unit{^\circ north}. To adapt this to a global model, we need to make a few additional assumptions. The tabulated height for each vegetation type $h(N)$ in the EMEP scheme is regarded as constant at mid latitudes (40--60\,\unit{^\circ}). Towards the poles, we decrease the height of each vegetation type using the same rate as described in \citet{ACP:Simpson2012}. At a latitude of 74\,\unit{^\circ} a minimum height of $3/10 \cdot h(N)$ is reached and kept constant. Towards the equator, we increase the height linearly so that at a latitude of 10\,\unit{^\circ} a maximum height of $2 \cdot h(N)$ is achieved which is then held constant. We also assume symmetry in both hemispheres. Presuming a typical tree height of 20\,\unit{m} at mid latitudes, this step-wise function yields a height of 8\,\unit{m} at high latitudes and 40\,\unit{m} in the tropics which is not unrealistic. For four example PFTs, results are shown in the Supplement (Fig.~S.2).

@@

In agriculture, there are different empirical rules to estimate the SGS and EGS. The simplest assumption is that greening starts after 5 consecutive days with a daily average temperature above 5\,\unit{^\circ C} and vice versa for EGS. Based on this rule of thumb, we have pre-processed our meteorological input data offline. If, within these 5 days, the average temperature drops below / rises above 5\,\unit{^\circ C}, the counters are reset, respectively. First, we used the 5\,\unit{^\circ C}-days criteria for $45\,\unit{^\circ} < \text{lat} < 85\,\unit{^\circ}$ in the NH, but extended it also to $35\,\unit{^\circ} < \text{lat} < 65\,\unit{^\circ}$ in the SH. In all other cases and where the 5\,\unit{^\circ C}-days criteria fails, we use the SMOKE-BEIS parameterization. The described algorithm written in python 2.7 has been included as Supplement~S.4.1. An example map of the computed GLEN using the 5\,\unit{^\circ C}-days criteria in both hemispheres is shown in Fig.~\ref{fig:glen_2015_she}.
\begin{figure}[t]
  \includegraphics[width=8.3cm, clip, trim={0.cm 1.75cm 0.cm 1.75cm}]{pictures/GLEN_2005_she.pdf}
  \caption{Pre-processing of greening season from meteorological surface temperature fields. Shown is the total length of the greening season (GLEN) for the year 2005. The 5\,\unit{^\circ C}-days criteria has been used in both hemispheres mid--high latitudes. Ocean has been shaded to indicate that greening season will only affect land.}
  \label{fig:glen_2015_she}
\end{figure}
\subsubsection{Photosynthetic photon flux density}
From OpenIFS an accumulated PAR is available, but for practical use in Eq.~(\ref{eq:flight}) it needs to be deaccumulated. Since this field is already integrated assumingly over the spectrum $(400-700)\,\unit{nm}$, it is not PAR but PPFD. Therefor, we will refer to the deaccumulated field as PPFD. An example PAR for January 2nd 2005 is shown in Supplement~S.5.1. Deaccumulation of hour 3 till 21, simply means
  \text{PPFD}(t_i) = \text{PAR}(t_{i+1})-\text{PAR}(t_i).
If everything was right in the OpenIFS output, we would not need to do any further processing, however the first timestep of each day has already accumulated PAR from some other date and time in addition to the actual hour 0 data (see Supplement~S.5.2), so it is impossible to accurately deaccumulate this time step. The best choice is subtracting the difference between hour 21 and 12 of the previous day (Supplement~S.5.3)
  \text{PPFD}(t=0) = \text{PAR}(t=0) - \left[\text{PAR}(t=21\,\unit{h}-1\,\unit{day})-\text{PAR}(=12\,\unit{h}-1\,\unit{day})\right].
The resulting PPFD fields are still accumulated over a time period of three hours and should be devided by 3. A \href{https://confluence.ecmwf.int/display/CKB/ERA-Interim\%3A+surface+photosynthetically+active+radiation+\%28surface+PAR\%29+values+are+too+low}{known issue} in the OpenIFS (cycles $\le$ c41r2) causes surface PAR values to be about 30\,\unit{\%} below observations. To counter this, we decided to refrain from the devision at this stage.
\section{Evaluation}
\label{sec:eval}
In this section, we present results from a manifold of OsloCTM3 model runs testing different parameters of the EMEP scheme. We focus on the changes in ozone, e.g., total dry deposition, dry deposition velocities, and surface concentration, related to the stomatal conductance parameterization. We evaluate our results with respect to \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and observations (Section~\ref{subsec:obs}). The OsloCTM3 is driven by meteorological input fields from \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF -- OpenIFS version cy38r1nc4}. CEDS historical emission inventory is used for anthropogenic emissions, while biomass burning is covered by NASA's Global Fire Emissions Database, Version4 (GFEDv4). In the following, we will discuss the various model sensitivity studies.
\subsection{Sensitivity studies}
\label{subsec:sens}
As presented in Section~\ref{sec:model_des}, the OsloCTM3 dry deposition scheme hase been revised majorly. Due to the previously generalized implementation of the Wesely scheme with respect to dry deposition resistances and surface types, it is not possible to fully disentangle the effects of the changes. We have conducted 9 different model simulations to assess the parameter space of the EMEP scheme. A reference simulation featuring the Wesely scheme has been done and will be refered to as \emph{Wesely type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models. For all model integrations, the meteorological reference year is 2005. This choice only affects the comparison with data and multi-model studies that either perform analysis on decadal averages or differing years. After finishing the work on the integration of the EMEP parameterization and initial testing, \emph{EMEP\_full} was supposed to be the baseline simulation for our sensitivity studies, \emph{EMEP\_offLight} and \emph{EMEP\_offPhen}. As indicated by the names, these scenarios are extreme cases, setting $f_\text{light}$ and $f_\text{phen}$ to 1, respectively. During analysis of the results, we found, that the SMOKE-BEIS parameterization of the greening season did not extend to the boreal and subarctic regions. Especially in Europe, basically resulting in the subpression of canopy resistance in these regions. The following sensitivity study therefor comprises two different versions of pre-processed greening season (refer to Section~\ref{subsec:pre-pro}), mid and high latitudes in the borthern hemisphere only (\emph{EMEP\_ppgs}) and for both, northern and southern hemisphere (\emph{EMEP\_ppgssh}). Building on the latter, we want to confirm the importance of the $v^\chem{O_3}_\text{ice/snow}$ update \citep{ACP:Helmig2007} in the Subarctic and Arctics, within the framework of the OsloCTM3.

focussion on the details of the canopy resistance parameterization. 
 For simplicity, we only list the tested parameters. An x in the table denotes that the model was run exactly in the configuration as has been decribed in Section~\ref{sec:model_des}. With respect to the sensitivity on light (PPFD), $f_\text{light}\equiv 1$ has been tested. Regarding the phenology, we have tested three different options: No phenology ($f_\text{phen}\equiv 1$), phenology including / excluding the Tropics (TR). Diverging from \citet{ACP:Simpson2012}, the simulations inititally ran with the uppermost soil water level (SWVL1). One sensitivity test was conducted using SWVL4. The meteorological year used in all simulations is 2005. Most of the simulations were run with CEDS emissions for 2014. One test has been conducted to look at the impact of the emissions. All simulations discussed in this section are summarized in Table~\ref{tab:simsum}.
\begin{table*}[t]
  \caption{Summary of specifications of all simulations discussed in this section. For simplicity, only the tested parameters are listed. An x denotes that the model was run exactly in the configuration as has been decribed in Section~\ref{sec:model_des}.}
  \begin{tabular}{lccccccccc}
    \tophline
    Simulation & Wesely scheme & \multicolumn{3}{c}{EMEP scheme} & \multicolumn{2}{c}{Greening season} & $v^\chem{O_3}_\text{ice/snow}$ & Forcings\\
    & & $f_\text{phen}$ & $f_\text{light}$ & $f_\text{SW}$ & SMOKE-BEIS & 5\,\unit{^\circ C}-days & (\unit{cm\,s^{-1}}) & (year)\\
    \middlehline
    Wesely type       & x & \multicolumn{3}{c}{n/a}  & x &       & 1/2000  & 2014 \\
    EMEP\_full        &   & incl. TR & x & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_offLight    &   & incl. TR & 1 & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_offPhen     &   & 1        & x & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_SWVL4       &   & incl. TR & x & SWLV4     & x &       & 1/2000  & 2014 \\
    EMEP\_ppgs        &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2014 \\
    EMEP\_ppgs\_2005  &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2005 \\
    EMEP\_ppgssh      &   & excl. TR & x & SWLV1     &   & NH+SH & 1/2000  & 2014 \\
    EMEP\_ppgssh\_ice &   & excl. TR & x & SWLV1     &   & NH+SH & 1/10000 & 2014 \\
    \bottomhline
  \end{tabular}
  \label{tab:simsum}
  %\belowtable{} % Table Footnotes
\end{table*}
\citep[e.g., $v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{cm\,s^{-1}}$,][]{ACP:Helmig2007}
dry dep velocity over water \citep{JGR:Helmig2012} Impact of ocean? Although it is a low number there is 2\/3 ocean on earth... (Think, I read this in Hardacre paper... -> Luhar)
